---
title: "Doing Data Science Case Study 1"
author: "Benjamin Wilke"
date: "February 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load raw data: Beers.csv and Breweries.csv

```{r echo=TRUE}
beers <- read.csv("Beers.csv", header = TRUE)
breweries <- read.csv("Breweries.csv", header = TRUE)
```

How many breweries are present in each state?

```{r echo=TRUE}
library(plyr)
count(breweries, 'State')
```

Merge beer data with breweries data.

```{r echo=TRUE}
head(beers)
head(breweries)
colnames(breweries)[1] <- "Brewery_id"

breweriesbeers <- merge(breweries, beers, by = "Brewery_id")
colnames(breweriesbeers)

colnames(breweriesbeers)[2] <- "BreweryName"
colnames(breweriesbeers)[5] <- "BeerName"

breweriesbeers <- breweriesbeers[,c(1, 6, 2, 3, 4, 5, 7, 8, 9)]
colnames(breweriesbeers)

head(breweriesbeers)
tail(breweriesbeers)
```

Report the number of NAs in each column.

```{r echo=TRUE}
sapply(breweriesbeers, function(x) sum(is.na(x)))
```

Compute the median alcohol content and IBU for each state. Plot a bar chart to compare.

```{r echo=TRUE}
ABVbyState <- data.frame(row.names = 1:51, tapply(breweriesbeers$ABV, breweriesbeers$State, function(x) median = median(x, na.rm = TRUE)))
IBUbyState <- data.frame(tapply(breweriesbeers$IBU, breweriesbeers$State, function(x) median = median(x, na.rm = TRUE)))

ABVbyState
IBUbyState

barplot(ABVbyState[2], xlab=ABVbyState[1])

```
IBU
Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?

```{r}
max(breweriesbeers$ABV, na.rm = TRUE)
breweriesbeers[which(breweriesbeers$ABV == 0.128), c("State", "ABV")]

max(breweriesbeers$IBU, na.rm = TRUE)
breweriesbeers[which(breweriesbeers$IBU == 138), c("State", "IBU")]

```

Summary statistics for the ABV variable.

```{r}
mean(breweriesbeers$ABV, na.rm = TRUE)
median(breweriesbeers$ABV, na.rm = TRUE)
sd(breweriesbeers$ABV, na.rm = TRUE)
```

Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.

```{r}

library(ggplot2)

ggplot(data = breweriesbeers) +
  geom_point(mapping = aes(x = IBU, y = ABV))
```
